On the properties of discs around accreting brown dwarfs

### 2010-07-16 BDD_submit/paper.tex
Various typos
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2383 25464f21-6099-402c-b30a-640f1d40d38b

Comparison of accretion rates across the pre-MS mass spectrum has
indicated that the accretion rate is strongly correlated with pre-MS
mass, \citep{muzerolle_2003,natta_2004,natta_2006}, with an
approximate form $\dot{M}_{\rm acc}\propto M^2_*$. Since in the
canonical picture the accretion rate is driven by the disc viscosity,
and should be independent of the mass of the central object, this
correlation is somewhat surprising. There is some danger that the
correlation is the result of, or at least strengthened by, the
presence of observational biases \citep{clarke_2006}. At the high-mass
(CTTS) end of the mass spectrum the lowest accretion-rates cannot be
measured via continuum methods since the excess is too small in
contrast with the photospheric emission. The emission lines will also
be weak, and indeed the H$\alpha$ equivalent width (EW) may be less
than the 10\AA\ that traditionally demarcates classical from
weak-lined T Tauri stars \citep[although this boundary may well be
spectral type dependent,][]{navascues_2003}. Such objects should show
doppler-broadened profiles, but the line wings will be weak, and the
presence of underlying H$\alpha$ absorption may become dominant. Thus
there may well be a population of low-accretion-rate CTTS which are
current missing from the surveys.

@@

In this study we assume that accretion from the central star occurs
along magnetically channelled columns from the inner disc boundary.
For CTTS stars, \cite{bouvier_2007} show that the magnetic truncation
radius ($R_{\rm mag}$) is less than the co-rotation radius ($R_{\rm
  co}$), where the angular Keplerian velocity of the disc is equal to
the surface angular velocity of the central star.  Calculations of the
magnetic truncation radius depend on derivations of the surface
magnetic field \citep{koenigl_1991}. This is currently unavailable for
BDs, due to an increased number of molecular species obscuring
the Zeeman splitting signatures that are normally used to derive
stellar surface magnetic fields. Therefore, for our model grid we have
adopted an initial inner disc radius as the co-rotation radius,
\begin{equation}
{R_{\rm inner}=\left({GM_*\tau^2\over{4\pi ^2}}\right)^{1\over{3}}},
\label{inner_eq}
\end{equation}
where $\tau$ is the stellar rotation period and $R_{\rm inner}$ is the
inner radius. This is effectively adopting a disc-locking mechanism
(without associated angular momentum loss), as for disc-locked stars,
$R_{\rm mag}\approx R_{\rm co}$,
\citep{koenigl_1991,shu_1994}. Therefore, for our
model simulations, this inner edge radius is dependent on, and derived
from, the adopted value of the rotational period for the central star,
as well as being weakly dependent on the stellar mass. As discussed in
Section \ref{intro}, the inner disc can be cleared through a number of
mechanisms, including binarity or giant planet formation,
photoevaporation or photoionisation of the disc and dust grain growth
or settling. For BDD systems where a disc is modeled a treatment of
dust sublimation is included (discussed in Section \ref{dust_edge}).
However, the effects of binarity or giant planet formation are
neglected. Further to the stellar mass and period required prior to
calculation of the inner disc radius, we require a disc mass (in
stellar masses).

@@

Little observational evidence can be found for approximate sizes of
accretion hot spots due to their more transient nature and often
smaller coverages, when compared to cooler or `plage' spots
\citep{herbst_2007}. \cite{bouvier_1995} modeled the
size of the cool spots on solar-type stars for a selection of
periodically variable candidates. They found projections of cooler
spots, onto the stellar disc, of a few to $\sim$60\%.
\cite{bouvier_1995} also found projected sizes, onto the
stellar disc, of typically a few \% to around 10\% for hot spots.
\cite{bertout_1996} used observations of YY Orionis monitoring
flux amplitude variations as a function of wavelength to derive a
probable hot spot area of around 10\%. The spot temperature was also
modeled for YY Orionis in \cite{bertout_1996}, resulting in a
best fitting areal coverage of 11\%. Therefore, to bound the probable
areal coverage range of the accretion hot spots we have adopted areal
coverages of 1 and 10\%.

@@

Previously studies modeling BD discs have adopted a range of disc mass
fractions, for instance \citep{walker_2004} use 0.1, 0.01 and
0.001$M_*$. \cite{wood_2002} fitted observed spectra with modeled SEDs
to derive a disc mass of 0.003$M_*$  for HH~30 IRS. Subsequent
derivations of disc masses have converged to within an order of
magnitude, with the following specific results: 0.03$M_*$
\citep[$\rho$ Ophiuchi,][]{natta_2002}, 0.055$M_*$ \citep[GM
Aurigae,][]{rice_2003}, 0.03$M_*$ \citep[GY 5, GY 11, and GY
310,][]{mohanty_2004} and 0.022$M_*$\citep[2MASS
J04442713+2512164,][]{bouy_2008}. As the derived disc masses all have
a similar order of magnitude we have adopted $M_{\rm disc}\approx
$\,0.01$M_*$. As changes in disc masses are expected to change the
resulting SED less than perhaps, accretion rate for example, we have
not varied the disc mass for this study. The results of simulations
varying this parameter will be published in a future paper.

@@

Figures \ref{age_mass_VVI} and \ref{age_mass_JJK} shows CMDs in $M_V$
\& $(V-I)_0$ and $M_J$ \& $(J-K)_0$ respectively. The left panels of
both figures shows stars classed as typical accretors with accretion
rates of \logmdot = $-$9, $-$10 and $-$11 \& $-$12, shown as blue,
black and red dots respectively in the top panels. The right panels of
Figures \ref{age_mass_VVI} and \ref{age_mass_JJK} show systems with
extreme accretion rates, with \logmdot = $-$6, $-$7 and $-$8 shown as
blue, black and red dots respectively, in the top right panel. The
bottom panels then show the systems separated into groups by
inclination. These groups are $\theta \leq 48^{\circ}$ as blue dots
(classed as face-on systems), $\theta> 56$ \& $64^{\circ}$ (classed as
the expected systems, as the expectation value of
$cos(\theta)=60^{\circ}$) as black dots and $\theta \geq 71^{\circ}$
(classed as edge-on systems) as red dots (for typical, bottom left,
and extreme, bottom right accretion rates). The left panels then show
naked stars isochrones (created from our grid of simulated photometry)
for 1 Myrs at accretion rates of \logmdot = $-$12 and $-$9, as solid
and dashed lines respectively (with coverage of 10\% and rotation
rates of 5 days). Whereas the left hand panels show naked stars
isochrones of \logmdot = $-$6 and $-$9 as solid and dashed black lines
respectively (with coverage of 10\% and rotation rates of 5 days. The
solid green line, in the top panels, shows the 1 Myr isochrone of
\cite{siess_2000} adjusted to a distance of 250 pc and an extinction of
$A_V=2$ mag, simulating a background population of CTTS stars.

@@

As can be seen in Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}
current accretion and disc presence in a star and disc systems creates
a scatter in our simulated photometry indicative of a much larger
isochronal age or mass spread. Indeed, for many BDD systems even at
nominal accretion rates of \logmdot = $-$11 or $-$12, for our
simulations, the colours of these stars move significantly blueward of
the expected BD locus in a $M_V$, $(V-I)_0$ CMD and redward in a
$M_J$, $(J-K)_0$ CMD. As, in the case of typical accretors, the input
variables are in the range expected for a BD population, one could
reasonably expect observed true BD loci to show a similar scatter. It
is clear that even a wide photometric selection would not even include
all of the negligibly accreting systems at expected
inclinations. Whilst the derivation of masses and ages for these
objects will be difficult. Furthermore, for the higher accretion rates
of \logmdot = $-$9 or $-$10 the movement of the star within the CMD
will effectively shift the star into the contamination region expected
for background CTTS or MS stars at a $(V-I)_0$ of $\leq$1.5 and $M_V$
12--10, and as such the star would not be included in a
photometrically selected BD sample. The solid green line, in the top
panels of Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}, showing
the 1 Myr isochrone of \cite{siess_2000} at a distance of 250 pc and
extinction of $A_V=2$ mags shows that the BDD systems with higher
accretion rates could easily be confused for a background CTTS or MS
population. Indeed misclassification of a BDD system as a CTTS system
has already been revealed in \cite{white_2003}. For the extreme
accretors, as shown in the right panels of Figures \ref{age_mass_VVI}
and \ref{age_mass_JJK}, there is little chance of these objects being
classed, photometrically, as BD candidates or being assigned the
correct mass. Therefore, if one attempted to locate a population of
BDD systems with elevated or extreme accretion rates, target selection
would have to be placed at much brighter magnitudes, and bluer for
optical or redder for IR, colours.

@@

Disc fractions have been derived using infrared excesses previously in
\textit{JHK} \citep[e.g][]{lada_1995,carpenter_1997}, however recent
works pre-dominantly use \textit{Spitzer} IRAC magnitudes
\citep[e.g][]{luhman_2005b,luhman_2008,gutermuth_2008,monin_2010}. Furthermore,
MIPS magnitudes are used to identify so-called debris discs, where IR
excesses are not apparent at shorter wavelengths
\citep[e.g][]{currie_2008,bryden_2009}. Finally, disc fractions have
also been derived using the $\alpha$ criteria, where
$\alpha=\frac{dlog\lambda F\lambda}{dlogF\lambda}$ between two
limiting wavelengths, originally used to distinguish amongst Class I,
II or II sources, but now used to detect disc presence
\citep{lada_2006,kennedy_2009}. An $\alpha>-2$ is used as a selection
criterion for disc presence for TTS stars. We have constructed the
$\alpha$ values for our model grid by adopting the limiting
wavelengths of \cite{kennedy_2009}, namely 3.6 to 8.0$\mu$m.

Figure \ref{disc_frac_JHJK} shows the photometry for all models in our
grid in a $(J-K)_0$, $(J-H)_0$ CoCoDs. In all panels all of the naked
systems are shown as black crosses. The left panels show those systems
with typical accretion rates and the right panels the extreme
accretors. The top panels of Figure \ref{disc_frac_JHJK} then separate
the systems by accretion rate with \logmdot = $-$9, $-$10 and $-$11 \&
$-$12, and $-$6, $-$7 and $-$8, shown as blue, black and red dots
respectively for the typical (top left panel) and extreme (top right
panel) accretors.  The bottom panels then shows the defined groups of
inclination angles, with $\theta \leq 48^{\circ}$ as blue dots,
$\theta> 56$ \& $64^{\circ}$ as black dots and $\theta \geq
71^{\circ}$ as red dots.

@@

Figure \ref{disc_frac_JHJK} shows that there is, for both typical and
extreme accretors, an overlap between the naked and BDD systems. This
suggests that an empirically placed cut in a CoCoD of this type (in
the absence of complications from variable reddening) could
mis-identify some candidates. However, as disc fractions defined by
placing a colour-colour cut are viewed as lower limits this effect may
be small. Our grid show no correlation in a CoCoD of this type with
rotation rate. In the case of the extreme accretors (right panels)
there is a strong correlation of accretion rate and scatter from the
naked stars. Additionally, the extreme accretors lie significantly
removed from the naked stars locus. There is a possibility that some of
these objects may be lost due to saturation and limiting magnitude
effects. Again, as with mass and age derivation, the present figures
are representative of similar figures using alternative, similar
magnitudes.

CoCoDs constructed using longer the longer wavelength bands of the
IRAC and MIPS cameras are most commonly used for disc fraction
calculation. Selection based on these types of data are well accepted
as indicators of disc presence. Figures \ref{disc_frac_irac} and
\ref{disc_frac_mips} show example CoCoDs for IRAC and MIPs magnitudes
respectively.

@@

Figure \ref{disc_frac_irac} shows that, as expected, the IRAC CoCoDs
clearly separate the naked and BDD systems. Once again, as found in
Figure \ref{disc_frac_JHJK}, for the extreme accretors the separation
of the BDD systems from the naked stars is a weak function of
accretion rate. There is also a weak correlation with inclination,
with the face-on and expected systems closer to the naked
stars. Figure \ref{disc_frac_irac} is representative of CoCoDs
constructed using other IRAC photometric channels. As we include
longer wavelength bands in the IRAC CoCoDs the separation between the
disc and naked loci increases. As the wavelength gets longer the flux
originates from regions of the disc at lower temperature and therefore
greater radial distance from the star. As photometric emission is
minimal past around 3\,$\mu$m, systems with discs will have
significantly different SEDs from naked systems.

@@

Recent derivations of disc fraction usually use colour-colour
selection in the IRAC photometric bands. As we have shown within our
model grid the separation between the BDD and naked systems is
clear. Therefore, we can examine the success of some recent
observationally placed when applied to our model grid. Additionally,
disc fractions have been derived using the $\alpha$ value
\citep{lada_2006}, essentially a slope of the SED between two
wavelengths (at wavelengths longer than the stellar flux peak). As
these $\alpha$ values are usually derived at wavelengths across the
IRAC bands one would expect the resulting disc fractions to be
reliable.

Figure \ref{disc_cuts} presents all the data for our model grid
separated by age, with 1 and 10 Myr BDD systems shown as blue and red
dots respectively. The naked systems are shown as black crosses. The
CoCoDs featured are from several recent publications where disc
fractions have been derived. The selection criteria from theses
studies are marked as dashed vertical and horizontal lines. The left
panels show the typical accretors and the right panels the extreme
accretors.

The observational cuts applied, shown as dashed lines, are from
\cite{luhman_2005b} a study of IC348, \cite{luhman_2008} a study of
$\sigma$ Orionis and \cite{monin_2010} \citep[using the criteria
of][]{gutermuth_2008} a study of the Taurus region. In these cases the
effects of extinction are either negligible in the plotted colours,
with values of $E([3.6]-[4.5])<0.04$ and $E([4.5]-[5.8])<0.02$ for
IC348 and $A_V \leq$4 mag \citep[which will be negligible in the IRAC
CoCoD,][]{allen_2004}, or the cuts have been placed in intrinsic
colour space as for $\sigma$ Orionis. The cuts from
\cite{luhman_2005b}, \cite{luhman_2008} and \cite{monin_2010} appear
as the top, top-middle and bottom-middle panels. The lower panel shows
a recent BDD candidate selection using the $\alpha$ value from
\citep{kennedy_2009}.

@@

Figure \ref{disc_cuts} shows that the recent observational cuts used
to identify disc candidates would be, in the main, reliable for our
model grid. Simulated photometry would be correctly identified using
these cuts. It is important to note that our conclusions so far have
been drawn from differential photometric arguments, in this case we
are using intrinsic colours and these values are extremely sensitive
to changes in zero point and photometric calibration. The top panel of
Figure \ref{disc_cuts} shows that other than a few typically accreting
systems the observational cut of \cite{luhman_2005b} would select all
of the BDD systems in our grid. The second panel down shows that the
cut of \cite{luhman_2008} would miss some typical and a very small
number of extreme accreting BDD systems (almost all are edge-on
systems). The third panel down shows only very few BDD systems will be
missed by the selection of \cite{gutermuth_2008} which was applied to
Taurus by\cite{monin_2010}. Finally, the selection for CTTS candidates
of $\alpha>-$2 \citep{kennedy_2009} appears reasonably applicable to
our BDD systems. As can be seen in the bottom panel of Figure
\ref{disc_cuts} almost all the typical, and all of the extreme
accreting BDD systems would be successfully identified using this
criterion. Again the missed typically accreting systems are edge-on
systems. Given, that extreme reddening may move naked stars into the
BDD region selected, in general, disc fractions are usually quoted as
lower limits. Therefore, a small number of miss-identified BDD system
is a negligible effect. This allows us to conclude, that ubiquitously
used BDD selections within IRAC CoCoDs are reliable when applied to
our model grid. Additionally, the $\alpha$ value would be a reliable
disc indicator for our model grid.

@@

We have constructed a model grid of SEDs, and subsequently photometric
magnitudes and colours, for actively accreting BDs with or without an
associated accretion disc. We have modeled the photospheric flux from
these BDs by adopting (and interpolating) the interior `DUSTY00'
models of \cite{chabrier_2000} combined with the `AMES-Dusty',
atmospheric models of \cite{chabrier_2000}. We have then assumed that
accretion occurs from an inner edge of a magnetically truncated
accretion disc (truncated at the co-rotation radius). The accretion
flux is calculated using a simple blackbody emission, given the
derivation of a characteristic spot effective temperature. SEDs were
then produced for both naked BDs and BDD systems. For the BDD systems
we have modeled the disc using the {\sc torus} radiative transfer code
using the \cite{lucy_1999} radiative equilibrium algorithm and
incorporating dust sublimation and including a treatment of vertical
hydrostatic equilibrium (see Section \ref{model} for a discussion of
the code). In order to produce a grid of simulated systems we have
varied several input parameters namely: stellar mass, stellar age,
stellar rotation rate, accretion rate, the areal coverage of the
accretion stream and the system inclination (the disc mass was
fixed). The ranges of these variables were selected to represent and
bound typical BD systems, justification is provided using evidence
from observational studies in Section \ref{par_space} and a final list
of the values of these variables can be found in Table
\ref{par_space_table}.

@@

Subsequently the SEDs of BDD systems with typical accretion rates and
associated discs change significantly from the assumed underlying
photospheric model flux, and therefore become difficult to
classify. In Section \ref{observables} we have shown that the BD
photosphere becomes veiled by the accretion flux for rates of
\logmdot$>-$9. The outer disc flaring observed in the BDD systems was
shown to cause occultation and a subsequent, sharp, fall in flux at an
inclination which decreases for more systems with higher accretion
rates. We have also shown that for extreme accretion rates the inner
wall increases in size and becomes convex in shape.

>>>>

Comparison of accretion rates across the pre-MS mass spectrum has
indicated that the accretion rate is strongly correlated with pre-MS
mass, \citep{muzerolle_2003,natta_2004,natta_2006}, with an
approximate form $\dot{M}\propto M^2_*$. Since in the
canonical picture the accretion rate is driven by the disc viscosity,
and should be independent of the mass of the central object, this
correlation is somewhat surprising. There is some danger that the
correlation is the result of, or at least strengthened by, the
presence of observational biases \citep{clarke_2006}. At the high-mass
(CTTS) end of the mass spectrum the lowest accretion-rates cannot be
measured via continuum methods since the excess is too small in
contrast with the photospheric emission. The emission lines will also
be weak, and indeed the H$\alpha$ equivalent width (EW) may be less
than the 10\AA\ that traditionally demarcates classical from
weak-lined T Tauri stars \citep[although this boundary may well be
spectral type dependent,][]{navascues_2003}. Such objects should show
doppler-broadened profiles, but the line wings will be weak, and the
presence of underlying H$\alpha$ absorption may become dominant. Thus
there may well be a population of low-accretion-rate CTTS which are
current missing from the surveys.

@@

In this study we assume that accretion from the central star occurs
along magnetically channelled columns from the inner disc boundary.
For CTTS stars, \cite{bouvier_2007} show that the magnetic truncation
radius ($R_{\rm mag}$) is less than the co-rotation radius ($R_{\rm
  co}$), where the angular Keplerian velocity of the disc is equal to
the surface angular velocity of the central star.  Calculations of the
magnetic truncation radius depend on derivations of the surface
magnetic field \citep{koenigl_1991}. This is currently unavailable for
BDs, due to an increased number of molecular species obscuring
the Zeeman splitting signatures that are normally used to derive
stellar surface magnetic fields. Therefore, for our model grid we have
adopted an initial inner disc radius as the co-rotation radius,
\begin{equation}
{R_{\rm inner}=\left({GM_*\tau^2\over{4\pi ^2}}\right)^{1\over{3}}},
\label{inner_eq}
\end{equation}
where $\tau$ is the stellar rotation period and $R_{\rm inner}$ is the
inner radius. This is effectively adopting a disc-locking mechanism
(without associated angular momentum loss), as for disc-locked stars,
$R_{\rm mag}\approx R_{\rm co}$,
\citep{koenigl_1991,shu_1994}. Therefore, for our model simulations,
this inner edge radius is dependent on, and derived from, the adopted
value of the rotational period for the central star, as well as being
weakly dependent on the stellar mass. As discussed in Section
\ref{intro}, the inner disc can be cleared through a number of
mechanisms, including binarity or giant planet formation,
photoevaporation or photoionisation of the disc and dust grain growth
or settling. For BDD systems where a disc is modelled a treatment of
dust sublimation is included (discussed in Section \ref{dust_edge}).
However, the effects of binarity or giant planet formation are
neglected. Further to the stellar mass and period required prior to
calculation of the inner disc radius, we require a disc mass (in
stellar masses).

@@

Little observational evidence can be found for approximate sizes of
accretion hot spots due to their more transient nature and often
smaller coverages, when compared to cooler or `plage' spots
\citep{herbst_2007}. \cite{bouvier_1995} modelled the
size of the cool spots on solar-type stars for a selection of
periodically variable candidates. They found projections of cooler
spots, onto the stellar disc, of a few to $\sim$60\%.
\cite{bouvier_1995} also found projected sizes, onto the
stellar disc, of typically a few \% to around 10\% for hot spots.
\cite{bertout_1996} used observations of YY Orionis monitoring
flux amplitude variations as a function of wavelength to derive a
probable hot spot area of around 10\%. The spot temperature was also
modelled for YY Orionis in \cite{bertout_1996}, resulting in a
best fitting areal coverage of 11\%. Therefore, to bound the probable
areal coverage range of the accretion hot spots we have adopted areal
coverages of 1 and 10\%.

@@

Previously studies modeling BD discs have adopted a range of disc mass
fractions, for instance \citep{walker_2004} use 0.1, 0.01 and
0.001$M_*$. \cite{wood_2002} fitted observed spectra with modelled SEDs
to derive a disc mass of 0.003$M_*$  for HH~30 IRS. Subsequent
derivations of disc masses have converged to within an order of
magnitude, with the following specific results: 0.03$M_*$
\citep[$\rho$ Ophiuchi,][]{natta_2002}, 0.055$M_*$ \citep[GM
Aurigae,][]{rice_2003}, 0.03$M_*$ \citep[GY 5, GY 11, and GY
310,][]{mohanty_2004} and 0.022$M_*$\citep[2MASS
J04442713+2512164,][]{bouy_2008}. As the derived disc masses all have
a similar order of magnitude we have adopted $M_{\rm disc}\approx
$\,0.01$M_*$. As changes in disc masses are expected to change the
resulting SED less than perhaps, accretion rate for example, we have
not varied the disc mass for this study. The results of simulations
varying this parameter will be published in a future paper.

@@

Figures \ref{age_mass_VVI} and \ref{age_mass_JJK} shows CMDs in $M_V$
\& $(V-I)_0$ and $M_J$ \& $(J-K)_0$ respectively. The left panels of
both figures shows stars classed as typical accretors with accretion
rates of \logmdot = $-$9, $-$10 and $-$11 \& $-$12, shown as blue,
black and red dots respectively in the top panels. The right panels of
Figures \ref{age_mass_VVI} and \ref{age_mass_JJK} show systems with
extreme accretion rates, with \logmdot = $-$6, $-$7 and $-$8 shown as
blue, black and red dots respectively, in the top right panel. The
bottom panels then show the systems separated into groups by
inclination. These groups are $\theta \leq 48^{\circ}$ as blue dots
(classed as face-on systems), $\theta> 56$ \& $64^{\circ}$ (classed as
the expected systems, as the expectation value of
$cos(\theta)=60^{\circ}$) as black dots and $\theta \geq 71^{\circ}$
(classed as edge-on systems) as red dots (for typical, bottom left,
and extreme, bottom right, accretion rates). The left panels then show
naked stars isochrones (created from our grid of simulated photometry)
for 1 Myrs at accretion rates of \logmdot = $-$12 and $-$9, as solid
and dashed lines respectively (with an areal coverage of 10\% and
rotation rates of 5 days). Whereas the left hand panels show naked
star isochrones of \logmdot = $-$6 and $-$9 as solid and dashed black
lines respectively (with an areal coverage of 10\% and rotation rates
of 5 days). The solid green line, in the top panels, shows the 1 Myr
isochrone of \cite{siess_2000} adjusted to a distance of 250 pc and an
extinction of $A_V=2$ mag, simulating a background population of CTTS
stars.

@@

As can be seen in Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}
current accretion and disc presence in a star and disc systems creates
a scatter in our simulated photometry indicative of a large isochronal
age or mass spread. Indeed, for many BDD systems even at nominal
accretion rates of \logmdot = $-$11 or $-$12, for our simulations, the
colours of these stars move significantly blueward of the expected BD
locus in a $M_V$, $(V-I)_0$ CMD (and redward in a $M_J$, $(J-K)_0$
CMD). As, in the case of typical accretors, the input variables are in
the range expected for a BD population, one could reasonably expect
observed true BD loci to show a similar scatter. It is clear that even
a wide photometric selection would not even include all of the
negligibly accreting systems at expected inclinations. Furthermore,
the derivation of masses and ages for these objects will be
difficult. Additionally, for the higher accretion rates of \logmdot =
$-$9 or $-$10 the movement of the star within the CMD will effectively
shift the star into the contamination region expected for background
CTTS or MS stars at a $(V-I)_0$ of $\leq$1.5 and $M_V$ 12--10, and as
such the star would not be included in a photometrically selected BD
sample. The solid green line, in the top panels of Figures
\ref{age_mass_VVI} and \ref{age_mass_JJK}, showing the 1 Myr isochrone
of \cite{siess_2000} at a distance of 250 pc and extinction of $A_V=2$
mags, shows that the BDD systems with higher accretion rates could
easily be confused for a background CTTS or MS population. Indeed
misclassification of a BDD system as a CTTS system has already been
revealed in \cite{white_2003}. For the extreme accretors, as shown in
the right panels of Figures \ref{age_mass_VVI} and \ref{age_mass_JJK},
there is little chance of these objects being classed,
photometrically, as BD candidates or being assigned the correct
mass. Therefore, if one attempted to locate a population of BDD
systems with elevated or extreme accretion rates, target selection
would have to be placed at much brighter magnitudes, and bluer for
optical or redder for IR, colours.

@@

Disc fractions have been derived using infrared excesses previously in
\textit{JHK} \citep[e.g][]{lada_1995,carpenter_1997}, however recent
works pre-dominantly use \textit{Spitzer} IRAC magnitudes
\citep[e.g][]{luhman_2005b,luhman_2008,gutermuth_2008,monin_2010}. Furthermore,
MIPS magnitudes are used to identify so-called debris discs, where IR
excesses are not apparent at shorter wavelengths
\citep[e.g][]{currie_2008,bryden_2009}. Finally, disc fractions have
also been derived using the $\alpha$ criteria, where
$\alpha=\frac{dlog\lambda F\lambda}{dlogF\lambda}$ between two
limiting wavelengths, originally used to distinguish amongst Class I,
II or III sources, but now used to detect disc presence
\citep{lada_2006,kennedy_2009}. An $\alpha>-2$ is used as a selection
criterion for disc presence for TTS stars. We have constructed the
$\alpha$ values for our model grid by adopting the limiting
wavelengths of \cite{kennedy_2009}, namely 3.6 to 8.0$\mu$m.

Figure \ref{disc_frac_JHJK} shows the photometry for all models in our
grid in a $(J-K)_0$, $(J-H)_0$ CoCoDs. In all panels, all of the naked
systems are shown as black crosses. The left panels show those systems
with typical accretion rates and the right panels the extreme
accretors. The top panels of Figure \ref{disc_frac_JHJK} then separate
the systems by accretion rate with \logmdot = $-$9, $-$10 and $-$11 \&
$-$12, and $-$6, $-$7 and $-$8, shown as blue, black and red dots
respectively for the typical (top left panel) and extreme (top right
panel) accretors.  The bottom panels then shows the defined groups of
inclination angles, with $\theta \leq 48^{\circ}$ as blue dots,
$\theta> 56$ \& $64^{\circ}$ as black dots and $\theta \geq
71^{\circ}$ as red dots.

@@

Figure \ref{disc_frac_JHJK} shows that there is, for both typical and
extreme accretors, an overlap between the naked and BDD systems. This
suggests that an empirically placed cut in a CoCoD of this type (in
the absence of complications from variable reddening) could
mis-identify some candidates. However, as disc fractions defined by
placing a colour-colour cut are viewed as lower limits this effect may
be small. The photometry from our grid shows no correlation in a CoCoD
of this type with rotation rate. In the case of the extreme accretors
(right panels) there is a strong correlation of accretion rate and
scatter from the naked stars. Additionally, the extreme accretors lie
significantly removed from the naked stars locus. There is a
possibility that some of these objects may be lost due to saturation
and limiting magnitude effects. Again, as with mass and age
derivation, the present figures are representative of similar figures
using alternative, passbands within the instrument filter sets.

CoCoDs constructed using the longer wavelength bands of the IRAC and
MIPS cameras are most commonly used for disc fraction
calculation. Selection based on these types of data are well accepted
as indicators of disc presence. Figures \ref{disc_frac_irac} and
\ref{disc_frac_mips} show example CoCoDs for IRAC and MIPs magnitudes
respectively.

@@

Figure \ref{disc_frac_irac} shows that, as expected, the IRAC CoCoDs
clearly separate the naked and BDD systems. Once again, as found in
Figure \ref{disc_frac_JHJK}, for the extreme accretors the separation
of the BDD systems from the naked stars is a weak function of
accretion rate. There is also a weak correlation with inclination,
with the face-on and expected systems closer to the naked
stars. Figure \ref{disc_frac_irac} is representative of CoCoDs
constructed using other IRAC photometric channels. As we include
longer wavelength bands in the IRAC CoCoDs the separation between the
disc and naked loci increases. As the wavelength gets longer the flux
originates from regions of the disc at lower temperature and therefore
greater radial distance from the star. As photometric emission is
minimal past around 3\,$\mu$m, systems with discs will have
significantly different SEDs from naked systems at these wavelengths.

@@

Recent derivations of disc fractions usually use colour-colour
selection in the IRAC photometric bands. As we have shown within our
model grid the separation between the BDD and naked systems is
clear. Therefore, we can examine the success of some recent
observationally placed cuts when applied to our model
grid. Additionally, disc fractions have been derived using the
$\alpha$ value \citep{lada_2006}, essentially a slope of the SED
between two wavelengths (at wavelengths longer than the stellar flux
peak). As these $\alpha$ values are usually derived at wavelengths
across the IRAC bands one would expect the resulting disc fractions to
be reliable.

Figure \ref{disc_cuts} presents all the data for our model grid
separated by age, with 1 and 10 Myr BDD systems shown as blue and red
dots respectively. The naked systems are shown as black crosses. The
colour combinations used to construct the CoCoDs featured are from
several recent publications where disc fractions have been
derived. The selection criteria from theses studies are marked as
dashed lines. The left panels show the typical accretors and the right
panels the extreme accretors.

The observational cuts applied, shown as dashed lines, are from
\cite{luhman_2005b} a study of IC348, \cite{luhman_2008} a study of
$\sigma$ Orionis and \cite{monin_2010} a study of the Taurus region
\citep[using the criteria of][]{gutermuth_2008}. In these cases the
effects of extinction are either negligible in the plotted colours,
with values of $E([3.6]-[4.5])<0.04$ and $E([4.5]-[5.8])<0.02$ for
IC348 and $A_V \leq$4 mag \citep[which will be negligible in the IRAC
CoCoD,][]{allen_2004}, or the cuts have been placed in intrinsic
colour space as for $\sigma$ Orionis. The cuts from
\cite{luhman_2005b}, \cite{luhman_2008} and \cite{monin_2010} appear
as the top, top-middle and bottom-middle panels. The lower panel shows
a recent BDD candidate selection using the $\alpha$ value from
\citep{kennedy_2009}.

@@

Figure \ref{disc_cuts} shows that the recent observational cuts used
to identify disc candidates would be, in the main, reliable for our
model grid. Simulated photometry would be correctly identified using
these cuts. It is important to note that our conclusions so far have
been drawn from differential photometric arguments, in this case we
are using intrinsic colours and these values are extremely sensitive
to changes in zero point and photometric calibration. The top panel of
Figure \ref{disc_cuts} shows that other than a few typically accreting
systems the observational cut of \cite{luhman_2005b} would select all
of the BDD systems in our grid. The second panel down shows that the
cut of \cite{luhman_2008} would miss some typical and a very small
number of extreme accreting BDD systems (almost all are edge-on
systems). The third panel down shows only very few BDD systems will be
missed by the selection of \cite{gutermuth_2008} which was applied to
Taurus by \cite{monin_2010}. Finally, the selection for CTTS candidates
of $\alpha>-$2 \citep{kennedy_2009} appears reasonably applicable to
our BDD systems. As can be seen in the bottom panel of Figure
\ref{disc_cuts} almost all the typical, and all of the extreme
accreting BDD systems would be successfully identified using this
criterion. Again the missed typically accreting systems are edge-on
systems. Given, that extreme reddening may move naked stars into the
BDD region selected, in general, disc fractions are usually quoted as
lower limits. Therefore, a small number of miss-identified BDD system
is a negligible effect. This allows us to conclude, that ubiquitously
used BDD selections within IRAC CoCoDs are reliable when applied to
our model grid. Additionally, the $\alpha$ value would be a reliable
disc indicator for our model grid.

@@

We have constructed a model grid of SEDs, and subsequently photometric
magnitudes and colours, for actively accreting BDs with or without an
associated accretion disc. We have modelled the photospheric flux from
these BDs by adopting (and interpolating) the interior `DUSTY00'
models of \cite{chabrier_2000} combined with the `AMES-Dusty',
atmospheric models of \cite{chabrier_2000}. We have then assumed that
accretion occurs from an inner edge of a magnetically truncated
accretion disc (truncated at the co-rotation radius). The accretion
flux is calculated using a simple blackbody emission, given the
derivation of a characteristic spot effective temperature. SEDs were
then produced for both naked BDs and BDD systems. For the BDD systems
we have modelled the disc using the {\sc torus} radiative transfer
code. The {\sc torus} code implements the \cite{lucy_1999} radiative
equilibrium algorithm, incorporates dust sublimation and includes a
treatment of vertical hydrostatic equilibrium (see Section \ref{model}
for a discussion of the code). In order to produce a grid of simulated
systems we have varied several input parameters namely: stellar mass,
stellar age, stellar rotation rate, accretion rate, the areal coverage
of the accretion stream and the system inclination (the disc mass was
fixed). The ranges of these variables were selected to represent and
bound typical BD systems, justification is provided using evidence
from observational studies in Section \ref{par_space} and a final list
of the values of these variables can be found in Table
\ref{par_space_table}.

@@

Subsequently the SEDs of BDD systems with typical accretion rates and
associated discs change significantly from the assumed underlying
photospheric model flux, and therefore become difficult to
classify. In Section \ref{observables} we have shown that the BD
photosphere becomes veiled by the accretion flux for rates of
\logmdot$>-$9. The outer disc flaring observed in the BDD systems was
shown to cause occultation and a subsequent, sharp, fall in flux at an
inclination which decreases for systems with higher accretion
rates. We have also shown that for extreme accretion rates the inner
wall increases in size and becomes convex in shape.
